---
title: "Gompertz Mixture fit of lifespan of WT BY4742, BY4741 genetic backgrounds and single deletion mutants"
author: "Emine Guven"
date: "10 May 2021"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r}
library('varhandle')
library('flexsurv')
library('stringr')
library("base")
library("fitmix")
```


### load the lifespan data previously generated from "rls 2016-11-29.csv"
#####cooked lifespans previously using "rls 2016-08-26.csv" and write outputs to the "updatedRLS_05052021.csv"
```{r}
lifespan_data<-read.csv("updatedRLS_05052021.csv")
names(lifespan_data)
```
### mixture plot function
```{r mixture plot function}
plot.gompertz.components <- function(mixture,component.number,...) {
  curve(mixture$lambda[component.number] *
          dgompertz(x,shape=mixture$shape[component.number],
                    rate=mixture$rate[component.number]), add=TRUE, col="darkred")}
```             

### BY4742
```{r}
#BY4742 plots
WT.BY4742<- lifespan_data[lifespan_data$genotype=="BY4742",]
#no need to filter temp==30  since I did filtering in the beginning
#WT.BY4742.temp<-WT.BY4742[WT.BY4742$temp==30,]
WT.BY4742.media<-WT.BY4742[WT.BY4742$media=="YPD",]

WT.BY4742.media<-WT.BY4742.media[WT.BY4742.media$mat=="MATalpha",]

WT.BY4742.media= WT.BY4742.media[!is.na(WT.BY4742.media[,1]), ]
dim(WT.BY4742.media)

class(WT.BY4742.media$single_lifespan)
str(WT.BY4742.media$single_lifespan)
```

### BY4742 individuals
```{r BY4742}
BY4742_lifespan_list=list()

for (k in c(1:length(WT.BY4742.media$single_lifespan))){
 # k=c(94:95)
  fk<-WT.BY4742.media$single_lifespan[k]
  #f<-unfactor(fk[k])
  BY4742_lifespan_single<-as.numeric(unlist(str_split(fk, ",")))
  #BY4742_lifespan_list<-mean(ref_lifespan_single)
  BY4742_lifespan_list[[length(BY4742_lifespan_list)+1]]<-BY4742_lifespan_single
}

length(BY4742_lifespan_list)
cumulative_BY4742_lifespan<-unlist(BY4742_lifespan_list)
length(cumulative_BY4742_lifespan)
cumulative_BY4742_vector<-na.omit(cumulative_BY4742_lifespan)
length(cumulative_BY4742_vector)
```

### mixture BY4742 
```{r mixture BY4742 cumulative}
cumulative_BY4742_mixGompertz<-fitmixEM(cumulative_BY4742_vector,"gompertz", 2, initial=FALSE)

## original BY4742
print(cumulative_BY4742_mixGompertz$estimate)
print(cumulative_BY4742_mixGompertz$measures)
```
###BY4742 mixture plot
```{r BY4742 mixture plot}
dfBY4742=data.frame(lambda=c(cumulative_BY4742_mixGompertz$estimate[,1]),
              shape=c(cumulative_BY4742_mixGompertz$estimate[,2]),
              rate=c(cumulative_BY4742_mixGompertz$estimate[,3]))

plot(hist(cumulative_BY4742_vector),col="grey",border="white",freq=FALSE,ylim=c(0,0.08))
#pdf("plots/mixture_fits.pdf")
lines(density(cumulative_BY4742_vector),lty=2,col="blue")
sapply(1:2,plot.gompertz.components,mixture=dfBY4742)
legend("topright", c("lifespan", "pdf","gompertz.mixture"), 
       col=c("grey","blue","violet"), lwd=10,cex=0.75)
```

### BY4741
```{r BY4741}
WT.BY4741<- lifespan_data[lifespan_data$genotype=="BY4741",]
WT.BY4741.media<-WT.BY4741[WT.BY4741$media=="YPD",]
WT.BY4741.media<-WT.BY4741.media[WT.BY4741.media$mat=="MATa",]
WT.BY4741.media= WT.BY4741.media[!is.na(WT.BY4741.media[,1]), ]
dim(WT.BY4741.media)

class(WT.BY4741.media$single_lifespan)
str(WT.BY4741.media$single_lifespan)
```

```{r BY4741 individuals }
### BY4741 individuals 
BY4741_lifespan_list=list()

for (k in c(1:length(WT.BY4741.media$single_lifespan))){
  # k=c(94:95)
  fk<-WT.BY4741.media$single_lifespan[k]
  #f<-unfactor(fk[k])
  BY4741_lifespan_single<-as.numeric(unlist(str_split(fk, ",")))
  #BY4741_lifespan_list<-mean(ref_lifespan_single)
  BY4741_lifespan_list[[length(BY4741_lifespan_list)+1]]<-BY4741_lifespan_single
}

length(BY4741_lifespan_list)
cumulative_BY4741_lifespan<-unlist(BY4741_lifespan_list)
length(cumulative_BY4741_lifespan)

cumulative_BY4741_vector<-na.omit(cumulative_BY4741_lifespan)
length(cumulative_BY4741_vector)
```
### mixture BY4741
```{r mixture BY4741}
#no need to do sampling cause it is already small enough
cumulative_BY4741_mixGompertz<-fitmixEM(cumulative_BY4741_vector,"gompertz", 2, initial=FALSE)

print(cumulative_BY4741_mixGompertz$estimate)
print(cumulative_BY4741_mixGompertz$measures)
```

### mixture plot of BY4741
```{r mixture plot BY4741}

dfBY4741=data.frame(lambda=c(cumulative_BY4741_mixGompertz$estimate[,1]),
              shape=c(cumulative_BY4741_mixGompertz$estimate[,2]),
              rate=c(cumulative_BY4741_mixGompertz$estimate[,3]))

plot(hist(cumulative_BY4741_vector),xlab="lifespan,divisions",                 col="grey",border="white",ylim=c(0,0.07),freq=FALSE)
#pdf("plots/mixture_fitsBY4741.pdf")
lines(density(cumulative_BY4741_vector),lty=2,col="blue")
sapply(1:2,plot.gompertz.components,mixture=dfBY4741)
legend("topright", c("lifespan", "pdf","gompertz.mixture"), 
       col=c("grey","blue","violet"),
       lwd=10,cex=0.75)
#dev.off()
```            
      
### tor1
```{r tor1}
tor1<- lifespan_data[lifespan_data$genotype=="tor1",]
#WT.BY4741.temp<-WT.BY4741[WT.BY4741$temp==30,] 
#there is no need to this since I did filtering in the beginning
tor1.media<-tor1[tor1$media=="YPD",]
tor1.media<-tor1.media[tor1.media$mat=="MATalpha",]
tor1.media= tor1.media[!is.na(tor1.media[,1]), ]
dim(tor1.media)

class(tor1.media$single_lifespan)
str(tor1.media$single_lifespan)
```
```{r tor1 individuals }
### tor1 individuals 
tor1_lifespan_list=list()

for (k in c(1:length(tor1.media$single_lifespan))){
  # k=c(94:95)
  fk<-tor1.media$single_lifespan[k]
  #f<-unfactor(fk[k])
  tor1_lifespan_single<-as.numeric(unlist(str_split(fk, ",")))
  #tor1_lifespan_list<-mean(ref_lifespan_single)
  tor1_lifespan_list[[length(tor1_lifespan_list)+1]]<-tor1_lifespan_single
}

length(tor1_lifespan_list)
cumulative_tor1_lifespan<-unlist(tor1_lifespan_list)
length(cumulative_tor1_lifespan)

cumulative_tor1_vector<-na.omit(cumulative_tor1_lifespan)
length(cumulative_tor1_vector)
```
### mixture tor1
```{r mixture tor1}
#no need to do sampling cause it is already small enough
cumulative_tor1_mixGompertz<-fitmixEM(cumulative_tor1_vector,"gompertz", 2, initial=FALSE)

cumulative_tor1_mixGompertz$estimate
cumulative_tor1_mixGompertz$measures
```


###mixture plot of tor1

```{r mixture plot of tor1}
dftor1=data.frame(lambda=c(cumulative_tor1_mixGompertz$estimate[,1]),
              shape=c(cumulative_tor1_mixGompertz$estimate[,2]),
              rate=c(cumulative_tor1_mixGompertz$estimate[,3]))

plot(hist(cumulative_tor1_vector),xlab="lifespan,divisions",                 col="grey",border="white",freq=FALSE,ylim=c(0,0.04))
#pdf("plots/mixture_fitstor1.pdf")
lines(density(cumulative_tor1_vector),lty=2,col="blue")
sapply(1:2,plot.gompertz.components,mixture=dftor1)
legend("topright", c("lifespan", "pdf","gompertz.mixture"), 
       col=c("grey","blue","violet"),lwd=1,cex=0.75)
#dev.off()
```         

### fob1
```{r fob1}
fob1<- lifespan_data[lifespan_data$genotype=="fob1",]

#WT.BY4741.temp<-WT.BY4741[WT.BY4741$temp==30,] 
#there is no need to this since I did filtering in the beginning

fob1.media<-fob1[fob1$media=="YPD",]

fob1.media<-fob1.media[fob1.media$mat=="MATalpha",]

fob1.media= fob1.media[!is.na(fob1.media[,1]), ]
dim(fob1.media)

class(fob1.media$single_lifespan)
str(fob1.media$single_lifespan)
```
```{r fob1 individuals }
### fob1 individuals 
fob1_lifespan_list=list()

for (k in c(1:length(fob1.media$single_lifespan))){
  # k=c(94:95)
  fk<-fob1.media$single_lifespan[k]
  #f<-unfactor(fk[k])
  fob1_lifespan_single<-as.numeric(unlist(str_split(fk, ",")))
  #fob1_lifespan_list<-mean(ref_lifespan_single)
  fob1_lifespan_list[[length(fob1_lifespan_list)+1]]<-fob1_lifespan_single
}

length(fob1_lifespan_list)
cumulative_fob1_lifespan<-unlist(fob1_lifespan_list)
length(cumulative_fob1_lifespan)

cumulative_fob1_vector<-na.omit(cumulative_fob1_lifespan)
length(cumulative_fob1_vector)
```
### mixture fob1
```{r mixture fob1}
#no need to do sampling cause it is already small enough
cumulative_fob1_mixGompertz<-fitmixEM(cumulative_fob1_vector,"gompertz", 2, initial=FALSE)

cumulative_fob1_mixGompertz$estimate
cumulative_fob1_mixGompertz$measures
```


###mixture plot of fob1

```{r mixture plot of fob1}
dffob1=data.frame(lambda=c(cumulative_fob1_mixGompertz$estimate[,1]),
              shape=c(cumulative_fob1_mixGompertz$estimate[,2]),
              rate=c(cumulative_fob1_mixGompertz$estimate[,3]))

plot(hist(cumulative_fob1_vector),xlab="lifespan,divisions",                 col="grey",border="white",freq=FALSE,ylim=c(0,0.04))
#pdf("plots/mixture_fitsfob1.pdf")
lines(density(cumulative_fob1_vector),lty=2,col="blue")
sapply(1:2,plot.gompertz.components,mixture=dffob1)
legend("topright", c("lifespan", "pdf","gompertz.mixture"), 
       col=c("grey","blue","violet"),lwd=1,cex=0.75)
#dev.off()
```            
### sod2 lifespan
```{r}
sod2<- lifespan_data[lifespan_data$genotype=="sod2",]

#WT.BY4741.temp<-WT.BY4741[WT.BY4741$temp==30,] 
#there is no need to this since I did filtering in the beginning

sod2.media<-sod2[sod2$media=="YPD",]

sod2.media<-sod2.media[sod2.media$mat=="MATalpha",]

sod2.media= sod2.media[!is.na(sod2.media[,1]), ]
dim(sod2.media)

class(sod2.media$single_lifespan)
str(sod2.media$single_lifespan)

```
```
```{r sod2 individuals }
### sod2 individuals 
sod2_lifespan_list=list()

for (k in c(1:length(sod2.media$single_lifespan))){
  # k=c(94:95)
  fk<-sod2.media$single_lifespan[k]
  #f<-unfactor(fk[k])
  sod2_lifespan_single<-as.numeric(unlist(str_split(fk, ",")))
  #sod2_lifespan_list<-mean(ref_lifespan_single)
  sod2_lifespan_list[[length(sod2_lifespan_list)+1]]<-sod2_lifespan_single
}

length(sod2_lifespan_list)
cumulative_sod2_lifespan<-unlist(sod2_lifespan_list)
length(cumulative_sod2_lifespan)

cumulative_sod2_vector<-na.omit(cumulative_sod2_lifespan)
length(cumulative_sod2_vector)
```
### mixture sod2
```{r mixture sod2}
#no need to do sampling cause it is already small enough
cumulative_sod2_mixGompertz<-fitmixEM(cumulative_sod2_vector,"gompertz", 2, initial=FALSE)

cumulative_sod2_mixGompertz$estimate
cumulative_sod2_mixGompertz$measures
```

###mixture plot of sod2

```{r mixture plot of sod2}
dfsod2=data.frame(lambda=c(cumulative_sod2_mixGompertz$estimate[,1]),
              shape=c(cumulative_sod2_mixGompertz$estimate[,2]),
              rate=c(cumulative_sod2_mixGompertz$estimate[,3]))

plot(hist(cumulative_sod2_vector),xlab="lifespan,divisions",                 col="grey",border="white",freq=FALSE,ylim=c(0,0.04))
#pdf("plots/mixture_fitssod2.pdf")
lines(density(cumulative_sod2_vector),lty=2,col="blue")
sapply(1:2,plot.gompertz.components,mixture=dfsod2)
legend("topright", c("lifespan", "pdf","gompertz.mixture"), 
       col=c("grey","blue","violet"),lwd=1,cex=0.75)
#dev.off()
```            
### sir2 lifespan
```{r}
sir2<- lifespan_data[lifespan_data$genotype=="sir2",]
dim(sir2)
#WT.BY4741.temp<-WT.BY4741[WT.BY4741$temp==30,] 
#there is no need to this since I did filtering in the beginning

sir2.media<-sir2[sir2$media=="YPD",]
dim(sir2.media)

sir2.media<-sir2.media[sir2.media$mat=="MATalpha",]
dim(sir2.media)

sir2.media= sir2.media[!is.na(sir2.media[,1]), ]
dim(sir2.media)

```

### sch9 lifespan
```{r}
sch9<- lifespan_data[lifespan_data$genotype=="sch9",]

#WT.BY4741.temp<-WT.BY4741[WT.BY4741$temp==30,] 
#there is no need to this since I did filtering in the beginning

sch9.media<-sch9[sch9$media=="YPD",]

sch9.media<-sch9.media[sch9.media$mat=="MATalpha",]

sch9.media= sch9.media[!is.na(sch9.media[,1]), ]
dim(sch9.media)

class(sch9.media$single_lifespan)
str(sch9.media$single_lifespan)

```
```
```{r sch9 individuals }
### sch9 individuals 
sch9_lifespan_list=list()

for (k in c(1:length(sch9.media$single_lifespan))){
  # k=c(94:95)
  fk<-sch9.media$single_lifespan[k]
  #f<-unfactor(fk[k])
  sch9_lifespan_single<-as.numeric(unlist(str_split(fk, ",")))
  #sch9_lifespan_list<-mean(ref_lifespan_single)
  sch9_lifespan_list[[length(sch9_lifespan_list)+1]]<-sch9_lifespan_single
}

length(sch9_lifespan_list)
cumulative_sch9_lifespan<-unlist(sch9_lifespan_list)
length(cumulative_sch9_lifespan)

cumulative_sch9_vector<-na.omit(cumulative_sch9_lifespan)
length(cumulative_sch9_vector)
```
### mixture sch9
```{r mixture sch9}
#no need to do sampling cause it is already small enough
cumulative_sch9_mixGompertz<-fitmixEM(cumulative_sch9_vector,"gompertz", 2, initial=FALSE)

cumulative_sch9_mixGompertz$estimate
cumulative_sch9_mixGompertz$measures
```

###mixture plot of sch9

```{r mixture plot of sch9}
dfsch9=data.frame(lambda=c(cumulative_sch9_mixGompertz$estimate[,1]),
              shape=c(cumulative_sch9_mixGompertz$estimate[,2]),
              rate=c(cumulative_sch9_mixGompertz$estimate[,3]))

plot(hist(cumulative_sch9_vector),xlab="lifespan,divisions",                 col="grey",border="white",freq=FALSE,ylim=c(0,0.08))
#pdf("plots/mixture_fitssch9.pdf")
lines(density(cumulative_sch9_vector),lty=2,col="blue")
sapply(1:2,plot.gompertz.components,mixture=dfsch9)
legend("topright", c("lifespan", "pdf","gompertz.mixture"), 
       col=c("grey","blue","violet"),lwd=1,cex=0.75)
#dev.off()
```   

```{r, sir2}

sir2<- lifespan_data[lifespan_data$genotype=="sir2",]

sir2.media<-sir2[sir2$media=="YPD",]

sir2.media<-sir2.media[sir2.media$mat=="MATalpha",]

sir2.media= sir2.media[!is.na(sir2.media[,1]), ]
dim(sir2.media)

class(sir2.media$single_lifespan)
str(sir2.media$single_lifespan)




### sir2 individuals 
sir2_lifespan_list=list()

for (k in c(1:length(sir2.media$single_lifespan))){
  # k=c(94:95)
  fk<-sir2.media$single_lifespan[k]
  #f<-unfactor(fk[k])
  sir2_lifespan_single<-as.numeric(unlist(str_split(fk, ",")))
  #sir2_lifespan_list<-mean(ref_lifespan_single)
  sir2_lifespan_list[[length(sir2_lifespan_list)+1]]<-sir2_lifespan_single
}

length(sir2_lifespan_list)
cumulative_sir2_lifespan<-unlist(sir2_lifespan_list)
length(cumulative_sir2_lifespan)

cumulative_sir2_vector<-na.omit(cumulative_sir2_lifespan)
length(cumulative_sir2_vector)

sir2<-cumulative_sir2_vector

### mixture sir2
```
```{r}
#no need to do sampling cause it is already small enough
cumulative_sir2_mixGompertz<-fitmixEM(cumulative_sir2_vector,"gompertz", 2, initial=FALSE)


fitmixEM(sir2,"gompertz", 2, initial=FALSE)


```

```{r}
cumulative_sir2_mixGompertz$estimate
cumulative_sir2_mixGompertz$measures
cumulative_sir2_mixGompertz$cluster


###mixture plot of sch9


dfsir2=data.frame(lambda=c(cumulative_sir2_mixGompertz$estimate[,1]),
                  shape=c(cumulative_sir2_mixGompertz$estimate[,2]),
                  rate=c(cumulative_sir2_mixGompertz$estimate[,3]))

```

```{r}

hist(sir2,xlab="replicative lifespan, sir2",col="grey",border="white",breaks=5,las=1,
     freq=FALSE,ylim=c(0,0.08),ylab="probability")
lines(density(sir2),las=1,col="blue")
x<-sapply(1:2,plot.gompertz.components,mixture=dfsir2)
legend("topright", c("lifespan", "pdf", "gompertz.mixture"), 
       col=c("grey","blue","darkred"),lwd=1,cex=0.75)


```

```{r}

hist(cumulative_sir2_vector,breaks=5, xlab="replicative lifespan, divisions",col="grey",border="white",freq=FALSE,ylim=c(0,0.07))
#pdf("plots/mixture_fitssir2.pdf")
lines(density(cumulative_sir2_vector),lty=2,col="blue")
sapply(1:2,plot.gompertz.components,mixture=dfsir2)
legend("topright", c("lifespan", "pdf","gompertz.mixture"), 
       col=c("grey","blue","darkred"),lwd=1,cex=0.75)
#dev.off()


```





